---
title: "genthat experiments - error analysis"
output:
  html_document:
    df_print: paged
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

library(tidyverse)
library(pbapply)
library(ggplot2)
library(janitor)
library(stringr)
library(anytime)
library(DT)

# this is where the result of the parallel is stored
RUN_DIR=file.path("..", "runs/packages")
```

```{r aux funs}
package_stat <- function(path) {
  data.frame(
    on.exit_set_log=file.exists(file.path(path, "trace-on.exit--set", "parallel.log")),
    on.exit_set=file.exists(file.path(path, "trace-on.exit--set", "output", "all", "genthat-tracing.csv")),
    on.exit_seq=file.exists(file.path(path, "trace-on.exit--sequence", "output", "all", "genthat-tracing.csv")),
    onexit_set=file.exists(file.path(path, "trace-onexit--set", "output", "all", "genthat-tracing.csv")),
    onexit_seq=file.exists(file.path(path, "trace-onexit--sequence", "output", "all", "genthat-tracing.csv")),
    count_entry_seq=file.exists(file.path(path, "trace-count-entry--sequence", "output", "all", "genthat-tracing.csv")),
    count_exit_seq=file.exists(file.path(path, "trace-count-exit--sequence", "output", "all", "genthat-tracing.csv")),
    run_log=file.exists(file.path(path, "run", "parallel.log")),
    run=file.exists(file.path(path, "run", "output", "all", "genthat-runs.csv")),
    run_package=file.exists(file.path(path, "run-package", "output", "all", "genthat-tracing.csv")),
    coverage=file.exists(file.path(path, "coverage", "output", "all", "covr.RDS"))
  )
}

format_difftime <- function(x, ...) {
  if (is.na(x)) return(NA)

  x <- as.integer(x)
  sprintf(
    "%02d:%02d:%02d", 
    x %% (24*60*60) %/% (60*60),
    x %% (60*60) %/% 60,
    x %% 60 %/% 1
  )
}

format_size <- function(x) {
  if (is.na(x)) return(NA)
  
  units <- c("B", "kB", "MB", "GB", "TB", "PB")
  
  fmt <- function(x, i=1) {
    xx <- x / 1024
    if (abs(xx) > 1 && i < length(units)) {
      fmt(xx, i+1)
    } else {
      sprintf("%.2f %s", x, units[i])
    }
  }
  
  sapply(x, fmt, USE.NAMES=FALSE)
}

load_file_from_packages <- function(filepath, from, fun) {
  transmute(from, package, fname=file.path(path, filepath)) %>% 
    pmap_dfr(fun)
}

read_csv_checked <- function(...) {
  tryCatch(read_csv(...), error=function(e) {
    message(..1, ": error: ", e$message)
  }, warning=function(e) {
    message(..1, ": warning: ", e$message)
  })
}

load_csvs <- function(path, from, ...) {
  load_file_from_packages(path, from, function(package, fname) {
    df <- read_csv_checked(fname, ...)
    mutate(df, package=package) %>% select(package, everything())
  })
}

load_traces <- function(config) {
  load_csvs(file.path(str_c("trace-", config), "output", "all", "genthat-tracing.csv"), from=packages, 
    col_types=cols_only(
      file="c",
      output="c",
      error="c"
    )
  )
}
```

```{r parallel logs aux}
read_parallel_log <- function(package, fname) {
  read_tsv(
    fname, 
    col_types=cols_only(
      Seq="i",
      Starttime="d",
      JobRuntime="d",
      Exitval="i",
      Signal="i",
      Command="c"
    )
  ) %>%
  transmute(
    package=package,
    started=anytime(Starttime),
    ended=anytime(Starttime+JobRuntime),
    duration=ended-started,
    status=Exitval,
    signal=Signal,
    command=Command
  )
}

read_parallel_logs <- function(task_name, from) {
  load_file_from_packages(file.path(task_name, "parallel.log"), from=from, fun=read_parallel_log)
}

load_trace_log <- function(config, from) {
  read_parallel_logs(str_c("trace-", config), from=from) %>%
    mutate(
      config=str_replace(command, ".* --config ([^ ]+) .*", "\\1"),
      decorator=str_replace(config, "(.*)--(.*)", "\\1"),
      tracer=str_replace(config, "(.*)--(.*)", "\\2")
    ) %>%
    select(-config)
}
```

# Load data

```{r find packages}
all_packages <- 
  data_frame(path=list.dirs(RUN_DIR, recursive=FALSE)) %>%
  filter(!endsWith(path, "/1")) %>%
  mutate(package=basename(path))

all_packages <-
  bind_cols(all_packages, package_stat(all_packages$path))

all_packages_obs <-
  all_packages %>%
  gather(
    key="task", 
    value="success", 
    count_entry_seq, on.exit_set, run, run_package, coverage
  )

# the subset of packages that made it till the end
packages <- filter(
  all_packages, 
  on.exit_set==TRUE, 
#  on.exit_seq==TRUE, 
#  onexit_seq==TRUE, 
#  count_entry_seq==TRUE, 
  run==TRUE
)
```

```{r load data}
on_entry_seq <- load_traces("count-entry--sequence")
on_exit_seq <- load_traces("on.exit--sequence")

traces <- load_traces("on.exit--set")
runs <- load_csvs(file.path("run", "output", "all", "genthat-runs.csv"), from=packages, col_types=cols_only(
    file="c",
    test="c",
    nb="i",
    failed="i",
    error="l",
    warning="i",
    run_error="c",
    elapsed="d",
    output="c"
  )
)

trace_log <- load_trace_log("on.exit--set", from=filter(all_packages, on.exit_set_log))
run_log <- read_parallel_logs("run", filter(all_packages, run_log))
```

```{r process data}
sum_traces <- function(traces) {
  df <- 
    group_by(traces, package) %>% 
    summarise(traces=n(), tests=sum(is.na(error)), failures=sum(!is.na(error)))
  
  # a sanity check (the verbose code is for debugging the bad ones)
  stopifnot(
    mutate(df, d = traces - tests - failures) %>% 
      filter(d != 0) %>% 
      nrow() == 0
  )
  
  df
}

s_traces <- 
  sum_traces(traces) %>%
  filter(traces > 0) %>%
  mutate(tests_traces=tests / traces)

p_runs <-
  filter(runs, is.na(nb) | (nb - warning) <= 1) %>%
  transmute(
    package,
    fun=test,
    failed=ifelse(is.na(failed), FALSE, failed == 1),
    error=ifelse(is.na(error), FALSE, error),
    exception=ifelse(is.na(nb), TRUE, FALSE),
    passed=(failed + error + exception == 0),
    elapsed=ifelse(is.na(elapsed), 0, elapsed),
    file
  )

s_runs <-
  group_by(p_runs, package) %>%
  summarise(failed=sum(failed), error=sum(error), exception=sum(exception), passed=sum(passed))
```

## Summary

### Absolute

```{r}
n_traces <- sum(s_traces$traces)
n_tests <- sum(s_traces$tests)
n_runs <- sum(s_runs$passed)
```

Overall we have `r n_traces` traces out of which we generate `r n_tests` tests which is `r n_tests/n_traces*100`%.
We can run `r n_runs` tests which is `r n_runs/n_tests*100`% of tests and `r n_runs/n_traces*100` % of all traces.

### Relative

We tried to run `r nrow(all_packages)`. 
From that `r nrow(packages)` finished. 
From that `r nrow(s_traces)` had some traces and `r nrow(s_runs)` had some tests.

```{r}
summary_stats <-
  filter(s_traces, traces > 0) %>%
  left_join(s_runs, by="package") %>%
  mutate_all(funs(replace(., is.na(.), 0))) %>%
  transmute(package, passed_traces=passed / traces, passed_tests=ifelse(tests > 0, passed / tests, 0))
```

- Ratio of calls we can reprodue out of all captured calls

```{r}
summary(summary_stats$passed_traces)
```

- Ratio of calls we can reproduce out of the calls we can generate tests for

```{r}
summary(summary_stats$passed_tests)
```

## Participating packages

### Selected packages

```{r package list}
select(all_packages, package, path) %>% datatable()
```

### Failed packages

```{r}
all_packages_obs %>%
  filter(!success) %>%
  count(task) %>%
  datatable()
```

```{r fig.height=40}
all_packages_obs %>%
  filter(!(task %in% c("onexit_set", "count_exit_seq", "coverage", "run_package")), success==FALSE) %>%
  ggplot(aes(
    x=factor(
      task, 
      levels=c(
        "count_entry_seq", 
        "count_exit_seq", 
        "on.exit_set", 
        "on.exit_seq", 
        "onexit_set", 
        "onexit_seq", 
        "run", 
        "run_package", 
        "coverage"
      )
    ), 
    y=package)
  ) + 
  geom_point() + 
  theme(strip.text.y=element_text(angle = 0)) +
  labs(
    title="Running failures",
    subtitle="Which packages failed on which tasks",
    x="Task",
    y="Package"
  )
```

- Failed tracing and generating

```{r}
select(trace_log, package, duration, status, signal) %>% filter(status != 0) %>% datatable()
```

- Failed running

```{r}
select(run_log, package, duration, status, signal) %>% filter(status != 0) %>% datatable()
```

# Results

## Errors

The following stats are for all the tests across all the packages.

### Error messages during tracing

```{r gen errors}
traces %>%
  mutate(error=ifelse(is.na(error), "No error", error)) %>%
  count(error) %>%
  arrange(desc(n)) %>%
  datatable()
```

### Categories of error messages during tracing

```{r}
traces %>%
  mutate(
    error=ifelse(is.na(error), "No error", error),
    error=ifelse(startsWith(error, "Generate error: Trace error:"), "Trace error", error),
    error=ifelse(startsWith(error, "Generate error"), "Generate error", error)
  ) %>%
  count(error) %>%
  arrange(desc(n)) %>%
  datatable()
```

### How much any of these contributes to tracing failures for individual packages

```{r}
n_success_traces <- nrow(filter(traces, is.na(error)))

traces %>%
  filter(!is.na(error)) %>%
  count(error) %>%
  filter(n > n_success_traces*0.001) %>%
  select(-n) %>%
  left_join(traces, by="error") %>%
  count(package, error) %>%
  left_join(s_traces, by="package") %>%
  mutate(r=n/traces, d=failures-n) %>%
  filter(r > .1) %>%
  arrange(desc(r)) %>%
  datatable()
```

#### Details for debugging

```{r}
generr_ratio <- function(pattern) {
  err <- traces %>% filter(grepl(pattern, error)) %>% count(package)
  s_traces %>% 
    right_join(err, by="package") %>% 
    mutate(r=n/traces, d=failures) %>% 
    arrange(desc(r))
}
```

- `Generate error: Serialization error: SEXP type S4SXP not supported!`

```{r}
generr_ratio("S4SXP") %>% filter(r > 0.1) %>% datatable()
```

- `Generate error: Serialization error: Serialized data structure contains cycle!`

```{r}
generr_ratio("Serialized data structure contains cycle") %>% filter(r > 0.1) %>% datatable()
```

- `Generate error: Serialization error: SEXP type PROMSXP not supported!`

```{r}
generr_ratio("PROMSXP") %>% filter(r > 0.1) %>% datatable()
```

- `Generate error: Serialization error: SEXP type EXTPTRSXP not supported!`

```{r}
generr_ratio("EXTPTRSXP") %>% filter(r > 0.1) %>% datatable()
```

- `Generate error: STRING_ELT() can only be applied to a 'character vector', not a 'NULL'`

```{r}
generr_ratio("STRING_ELT\\(\\) can only be applied to") %>% filter(r > 0.1) %>% datatable()
```

## Running

### Inconsitent data

Likely packages using testthat themselves

```{r}
runs %>% filter((nb - warning) > 1) %>% count(package) %>% datatable()
```

```{r run errors}
runs %>%
  transmute(run_error=ifelse(is.na(run_error), "No run_error", run_error)) %>%
  count(run_error) %>%
  arrange(desc(n)) %>%
  datatable()
```
