---
title: "R Notebook"
output: html_notebook
---

# Execution logs

In this section we look at the output form GNU parallel.

```{r}
read_parallel_log <- function(package, fname) {
  read_tsv(
    fname, 
    col_types=cols_only(
      Seq="i",
      Starttime="d",
      JobRuntime="d",
      Exitval="i",
      Signal="i",
      Command="c"
    )
  ) %>%
  transmute(
    package=package,
    started=anytime(Starttime),
    ended=anytime(Starttime+JobRuntime),
    duration=ended-started,
    status=Exitval,
    signal=Signal,
    command=Command
  )
}

read_parallel_logs <- function(task_name) {
  load_file_from_packages(file.path(task_name, "parallel.log"), read_parallel_log)
}
```

## Overall

```{r}

experiment <- read_parallel_log("all", "../runs/packages/parallel.log")
experiment %>% ggplot(aes(x="a", y=as.integer(duration, units="secs"))) + geom_boxplot()
```


## Tracing

### Running a package without genthat

```{r load run package parallel log}
run_package_log <- read_parallel_logs("run-package") %>%
  mutate(
    decorator=str_replace(command, ".* --config ([^ ]+) .*", "\\1")
  )
```

We have the run with two decorators, just to measure the overhead of the system.
The `none` decorator entirely skips using R site file so genthat is not touched.
There will still be a bit of an overhead, but similar should be expected when 
using `tools::testInstalledPackages`.

```{r}
prod_run_package_log <- run_package_log %>% filter(decorator=="none")
```

```{r fig.height=15}
prod_run_package_log %>%
  ggplot(aes(
    x=factor(package, levels=package[order(duration)]), 
    y=duration, 
    label=format_difftime(duration))
  ) +
  geom_col() +
  coord_flip() +
  scale_y_time() +
  geom_text(hjust=-.1, size=2) +
  labs(
    title="Package running time [decorator=none, tracer=NA]",
    subtitle="Running examples, tests, vignettes",
    x="Package",
    y="Time [hh:mm:ss]"
  )
```

### Running a package with genthat

```{r load trace parallel log}
trace_log <- read_parallel_logs("trace") %>%
  mutate(
    config=str_replace(command, ".* --config ([^ ]+) .*", "\\1"),
    decorator=str_replace(config, "(.*)--(.*)", "\\1"),
    tracer=str_replace(config, "(.*)--(.*)", "\\2")
  ) %>%
  select(-config)
```

Similarly here, we have multiple decorators and tracers.
For generating tests, we use `onexit` decorator and `set` tracer.
The others are just to find the missed opportunity used for analysis of the efectiveness.

```{r}
prod_trace_log <- trace_log %>% filter(decorator=="onexit", tracer=="set")
```


```{r package tracing time, fig.height=15}
prod_trace_log %>%
  ggplot(aes(
    x=factor(package, levels=package[order(duration)]), 
    y=duration, 
    label=format_difftime(duration))
  ) +
  geom_col() +
  coord_flip() +
  scale_y_time() +
  geom_text(hjust=-.1, size=2) +
  labs(
    title="Package tracing time [decorator=onexit, tracer=set]",
    subtitle="Running examples, tests, vignettes",
    x="Package",
    y="Time [hh:mm:ss]"
  )
```

```{r trace size, fig.height=10}
trace_size <- function(path) {
  rdss <- list.files(
    path=file.path(path, "trace", "output", "onexit--set", "all"), 
    pattern="\\.RDS$", 
    full.names=TRUE, 
    recursive=FALSE
  )
  
  sum(file.size(rdss))
}

traces_size <- result %>% mutate(size=sapply(path, trace_size))

traces_size %>%
  ggplot(
    aes(
      x=factor(package, levels=package[order(size)]), 
      y=size, 
      label=format_size(size)
    )
  ) +
  geom_col() +
  coord_flip() +
  geom_text(hjust=-.1, size=2) +
  # TODO: scale transfioramtion to show units
  labs(
    title="Tracing size [decorator=onexit, tracer=set]",
    subtitle="Running examples, tests, vignettes",
    x="Package",
    y="Size [bytes]"
  )
```

```{r fig.height=15}
prod_trace_log %>% 
  select(package, duration) %>%
  left_join(select(prod_run_package_log, package, duration), by="package") %>%
  mutate(diff=as.integer(duration.x)/as.integer(duration.y)) %>%
  ggplot(aes(x=factor(package, levels=package[order(diff)]), y=diff, label=sprintf("%.2fx", diff))) +
  geom_col() +
  geom_text(hjust=-.1, size=2) +
  coord_flip() +
  labs(
    title="Tracing vs Running",
    subtitle="How much is tracing slower",
    x="Package",
    y="Slowdown"
  )
```

```{r fig.height=10}
traces_size %>% 
  left_join(select(prod_trace_log, package, duration)) %>%
  ggplot(aes(x=as.integer(duration), y=size, label=package)) +
  geom_point() +
  geom_text(size=2, hjust=0.2, nudge_x=0.05) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title="Duration of tracing vs traces size",
    x="Duration (log)",
    y="Size (log)"
  )
```

### Summary

```{r fig.heigh=10}
prod_trace_log %>% 
  select(package, duration) %>%
  left_join(select(prod_run_package_log, package, duration), by="package") %>%
  transmute(tracing=as.integer(duration.x), running=as.integer(duration.y)) %>%
  gather(key="key", value="value", tracing, running) %>%
  ggplot(aes(x=key, y=value)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title="Summary of Tracing and Running",
    x="Task",
    y="Duration [s] (log)"
  )
```

```{r fig.height=10}
prod_trace_log %>% 
  select(package, duration) %>%
  left_join(select(prod_run_package_log, package, duration), by="package") %>%
  transmute(
    package, 
    tracing=as.integer(duration.x), 
    running=as.integer(duration.y),
    slowdown=tracing/running
  ) %>%
  ggplot(aes(x=running, y=tracing, label=package, color=slowdown)) +
  geom_jitter() +
  geom_text(size=2, hjust=-0.3, color="black") +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_gradient(low="green", high="red") + 
  labs(
    title="Running vs Tracing",
    x="Running time [s] (log)",
    y="Tracing time [s] (log)"
  )
```


## Test generation

```{r load generate parallel log}
generate_log <- read_parallel_logs("generate")
```

```{r test generating time, fig.height=15}
generate_log %>%
  ggplot(aes(
    x=factor(package, levels=package[order(duration)]), 
    y=duration, 
    label=format_difftime(duration))
  ) +
  geom_col() +
  coord_flip() +
  scale_y_time() +
  geom_text(hjust=-.1, size=2) +
  labs(
    title="Test generating time",
    x="Package",
    y="Time [hh:mm:ss]"
  )
```

## Test runnning

```{r load run parallel log}
run_log <- read_parallel_logs("run")
```

```{r test running time, fig.height=15}
run_log %>%
  ggplot(aes(
    x=factor(package, levels=package[order(duration)]), 
    y=duration, 
    label=format_difftime(duration))
  ) +
  geom_col() +
  coord_flip() +
  scale_y_time() +
  geom_text(hjust=-.1, size=2) +
  labs(
    title="Test running time",
    x="Package",
    y="Time [hh:mm:ss]"
  )
```

## Summary

### Genthat tests vs Package running time

```{r fig.height=15}
prod_run_package_log %>% 
  select(package, duration) %>%
  left_join(select(run_log, package, duration), by="package") %>%
  mutate(diff=as.integer(duration.y)/as.integer(duration.x)) %>%
  ggplot(aes(x=factor(package, levels=package[order(diff)]), y=diff, label=sprintf("%.2fx", diff))) +
  geom_col() +
  geom_text(hjust=-.1, size=2) +
  coord_flip() +
  labs(
    title="Genthat tests vs Package running time",
    subtitle="How much are the genthat tests slower",
    x="Package",
    y="Slowdown"
  )
```




